---
title: "Global Terrorism"
output:
  html_document: default
  html_notebook: default
---

```{r,warning=FALSE,message=FALSE}
#analyzing the GLobal Terrorist Activities
require(data.table)
require(readr) #to read data faster
require(ggplot2)
#Data preprocessing and Transformation packages
require(dplyr)
require(tidyr)


terror<-read_csv("F:/globalterrorismdb_0617dist.csv")

#checking the structure of data
#str(terror)
#summary(terror)
#a wide data set with lots of columns


#---------------------



```



Starting with Descriptive Analytics

```{r}
#checking how many terrorist attacks were successfull
table(terror$success)

success<-as.data.frame(table(terror$success))
success$Var1<-ifelse(success$Var1==1,"Successful","Unsuccessful")
names(success)<-c("Attack","count")
#almost 90% of all attacks were successfull
ggplot(aes(x = Attack,y=count), data = success) + 
  geom_col(width=0.5,color="black",fill="purple",alpha=0.5) + 
  labs(x = "Attack",y="Count",title="Distrubution of Successful and Unsuccessful attacks from 1970-2015")


```
Most of the attacks in the past were successfully executed.

---------------

####Analyzing what types of Attacks were done

Aggregating and summarising data.

```{r}
table(terror$attacktype1_txt)
#Most common terror attack is Bombing/Explosion

#Finding Succcessful attacks and their attack types
#percentage of Successful and unsuccessful attacks
#converting to a data frame
attackdf<-data.frame(table(terror$attacktype1_txt,
                                      terror$success))



#Spreading the data frame
attackdf<-attackdf %>% spread(key = Var2,Freq)
colnames(attackdf)<-c("Type_of_Attack","Unsuccessful","Successful")

#Adding new columns which contain Percentage of Successful and Unsucesful attacks
attackdf<-attackdf%>% mutate(PerSuccessful=round((Successful/(Unsuccessful+Successful))*100,2), PerUnsuccessful=round((Unsuccessful/(Unsuccessful+Successful))*100,2))

attackdf

#Plotting
plot<-ggplot(aes(x = "" , y = Successful,fill=Type_of_Attack),data = attackdf) + 
  geom_bar(width=1 , stat="identity") + 
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
        labs(title="Pie Chart of Count of the Types of Successful Attacks",
             x=NULL,y=NULL , fill = 'Type of Attack')


#Final Plot to make it a Pie Chart

plot + coord_polar(theta ="y" , start = 0)  + scale_color_brewer(palette = "Set1")



#Making a Bar plot
theme_set(theme_grey())

ggplot(aes(x = reorder(Type_of_Attack,PerSuccessful),y = PerSuccessful),data = attackdf) + 
  geom_col(color="black",fill="red",alpha=0.6) + 
  coord_flip() + 
  labs(title="Barplot of Types of Attacks and Successful attacks",x="Type of Attack",y="Percentage of Successful Attacks")
          




```

As the Plot shows the Most Successful *__Terror Attacks were for Hostage Taking,
Kidnapping,Armed Assaults, Bombings__*. 


----------------

####Analyzing the Countires and the Terrorist attacks took place 

```{r}
countrydf<-data.frame(table(terror$success,terror$country_txt))
#Spreading the Dataframe
countrydf<-countrydf %>% spread(Var1,Freq)
names(countrydf)<-c("Country","Unsuccessful","Successful")
countrydf<-countrydf %>% mutate(TotalAttacks=(Successful+Unsuccessful))
#renaming the columns


#FInding the Top countries with most succcessfully executed terror attacks

Topcountrydf<- countrydf %>% group_by(Country) %>%
        summarise(Mean_Successfull = mean(Successful)) %>%
        top_n(20) %>%
        arrange(desc(Mean_Successfull))


#Finding Countries with Most Unsuccessful Terror Attack attempts
TopFailsAttacks<- countrydf %>% group_by(Country) %>%
        summarise(Mean_Unsuccessfull = mean(Unsuccessful)) %>%
        top_n(20) %>%
        arrange(desc(Mean_Unsuccessfull))


#Countires with least Terrorist Attacks

LeastAttackdf<-countrydf %>% group_by(Country) %>%
      filter(TotalAttacks %in% seq(20,100)) %>%
      select(TotalAttacks) %>%
      top_n(30) %>%
      arrange(TotalAttacks)

#New Zealand ,Malaysia, Hong Kong , Cuba ,UAE etc are countries having very less
#Terrorist attacks less than 100 attempts.


#Plotting the Barplots
theme_set(theme_classic())
ggplot(aes(x = reorder(Country,Mean_Successfull) , y = Mean_Successfull),data = Topcountrydf) + 
  geom_col(color="black",fill="#1111F7") + 
  coord_flip() +
  labs(x = "Countries",y="Successfull Terrorist Attacks from 1970-2015")



#Countires with Most Unsuccessful Terror Attacks
theme_set(theme_bw())
ggplot(aes(x = reorder(Country,Mean_Unsuccessfull),y=Mean_Unsuccessfull),data = TopFailsAttacks) + 
  geom_col(stat="identity",color="black",fill="#DEFD16") + 
  coord_flip() +
  labs(x = "Countries",y="Unsuccessful Terrorist Attacks from 1970-2015") + 
  scale_y_continuous(limits = c(0,2100),breaks=seq(0,2100,200))



#Countries with Least Terrorist Activities 

leastFilter<- LeastAttackdf %>% 
  filter(TotalAttacks > 50) %>% top_n(20)
#Bar plot for Countries having attacks > 50
theme_set(theme_bw())
ggplot(aes(x = reorder(Country,-TotalAttacks),y=TotalAttacks),data = leastFilter) + 
  geom_col(stat="count",color="black",fill="#DEFD16") + 
  coord_flip() +
  labs(x = "Countries with Least Terrorist Attacks",y="Total Terrorist Attacks from 1970-2015")
  




```


----------


### What were the target of the Terrorists?


```{r}
worldTargetdf<-data.frame(table(terror$targtype1_txt)) %>% arrange(desc(Freq))

theme_set(theme_bw())
ggplot(aes(x = reorder(Var1,Freq),y = Freq),data = worldTargetdf) +
  geom_col(fill='purple') +
  coord_flip()




```

*__As we can notice from the plot above the most targetted entities were Citizens,then Military,Police ,Goverment and Businesses.__*


-----------

###Creating a new Dataframe

```{r}
countryList<-c("United States","United Kingdom")

SpecificTargdf<-terror %>% select(success,target1,targtype1_txt,attacktype1_txt,
                                  country_txt,city,iyear,gname,nkill) %>% 
                          filter(country_txt %in% countryList)
```

The above data frame consists only attacks on 2 major developed countries i.e __United States of America__ and __United Kingdom__ which were in the top 20 list of most successfully executed terrorist attacks.


---------------------



###Analyzing attack in USA

Creating a data frame with only some targetted variables.

```{r}
UStarg<-na.omit(SpecificTargdf) %>% filter(country_txt=="United States")
summary(UStarg$nkill)

```


####City with most successful attacks


```{r}
uscitydf<-data.frame(table(UStarg$city,UStarg$success))

uscitydf<-uscitydf %>% spread(Var2,Freq)
names(uscitydf)<-c("city",'unsuccessful','success')

uscitysuccess<-uscitydf %>% group_by(city) %>%
                          summarise(success=mean(success)) %>%
                          arrange(desc(success)) %>%
                          top_n(30)
#most attacks in New York
theme_set(theme_minimal())
ggplot(aes(x=reorder(city,success),y=success),data=uscitysuccess) +
  geom_col(fill="#131D75")+
  coord_flip() +
  labs(x="City",y="Number of successfull attacks")









```


####Time series of Year and number of succesful attacks in USA

Creating a dataframe with year and number of attacks.

```{r}

yearsuccess<-data.frame(table(UStarg$iyear,UStarg$success))
yearsuccess<-yearsuccess %>% spread(Var2,Freq)
names(yearsuccess)<-c("year","fail",'success')

#generating numeric year column
year<-data.frame(seq(1970,1992))
year[24:46,]<-seq(1994,2016)
names(year)<-c("year")

#adding the year df to yearsuccess df
yearsuccess$attack<-year






#Plotting the time series splot
success=ggplot(data=yearsuccess,aes(x=year,y=success)) +
  geom_point(color="purple",size=2) + 
  geom_line(color="#9124F2") + 
  #limits=c(1,46),breaks=seq(1,46,4)) 
  

fail=ggplot(data=yearsuccess,aes(x=as.numeric(year),y=fail)) +
  geom_point(color="purple",size=2) + 
  geom_line(color="#9124F2") + 
  scale_x_continuous(limits=c(1,46),breaks=seq(1,46,4))

success
fail


```








-----------------


###Analyzing attacks in India

I will create a new data frame for India by filtering and selecting only the relevant columns from the original data frame.

```{r}
#All the terrorist attacks in INDIA
indiaAttack<-terror %>%
  filter(country_txt=="India") %>% select(iyear,provstate,city,summary,location,success,attacktype1_txt,targtype1_txt,gname,weaptype1_txt,nkill) 

#Which state has most attacks 
table(indiaAttack$provstate)
indiastate<-data.frame(table(indiaAttack$provstate,
                             indiaAttack$success))

indiastate<-indiastate %>% spread(Var2,Freq)
names(indiastate)<-c("state","Unsuccessful","successful")

#plotting states and number of Successfull attacks
ggplot(aes(x=reorder(state,successful),y=successful ),data = indiastate) + geom_col(color="black",fill="blue",alpha=0.6) + 
  coord_flip() + 
  scale_y_continuous(limits=c(0,2000),breaks=seq(0,2000,400)) + 
  labs(x = "States",y="Successful terrorist attacks")
#most successfull terrorist attacks in Jammu and Kashmir
#We have Punjab , J and K and Assam on the top with most Successfull attacks
  
#number of kills
ggplot(aes(x = reorder(provstate,nkill), y = nkill),data = na.omit(indiaAttack)) + geom_col(fill="#E63B10") + 
  coord_flip() +
  labs(x="State",y="Number of Kills from 1970-2015")



ggplot(aes(x = reorder(provstate,nkill)),data = indiaAttack) + 
  geom_bar(fill="green",color="black") + 
  coord_flip()
  

```

We have Punjab , J & K and Assam on the top with most Successful terrorist attacks and maximum number of fatalities due to the terrorist activities occurred at Chattisgarh,Assam and J&K as well.

```{r}
#cities with most terrorist attacks
citydf<-data.frame(table(indiaAttack$city,indiaAttack$success))
citydf<-citydf %>%spread(Var2,Freq)
names(citydf)<-c("city","unsuccess","success")

#dataframe consisting of cities with top most count of successful attacks
cityTopsuccess<-citydf %>% group_by(city) %>%
  summarise(success = mean(success)) %>%
  arrange(desc(success)) %>%
  top_n(25)

#Plot for cities with most successful attacks added
ggplot(aes(x = reorder(city,success) , y = success),data = cityTopsuccess) + 
  geom_col(fill="#FE3C01") + 
  coord_flip() + 
  scale_y_continuous(limits=c(0,600),breaks=seq(0,600,100)) +
  labs(x="city name",y="Count of Successful terrorist attacks")


```

-----------------------


####Checking the type of attacks occured in India


```{r}
india_typeAttack<-with(indiaAttack,data.frame(table(attacktype1_txt,success)))
#spreading the dataframe
india_typeAttack<-india_typeAttack %>% spread(success,Freq)
india_typeAttack
#Most Bombings occured in india followed by armed assaults.
names(india_typeAttack)<-c("type","fail","success")

india_typeAttack<-india_typeAttack%>% mutate(successRate=round((success/(success+fail))*100,2) , failRate=round((fail/(success+fail))*100,2))

#plotting barplots now
ggplot(aes(y = successRate , x = reorder(type,successRate)),data = india_typeAttack) + geom_col(width=0.8,fill='#0000A4',alpha=0.7) +
  coord_flip() +
  labs(x="Type of attack",y="Success Rate of Attacks(in %)")




```

The surprising thing is that the __Bombings/Explosions__ have the highest distribution i.e highest count amongst all other types of terror attacks but their success rate is less.The terrorist activity having highest success rate is __unarmed assaults__.

---------------------------


#### Analyzing the Target of the Terrorist activities


```{r}

target_india<-with(indiaAttack,data.frame(table(targtype1_txt,success)))
target_india<-target_india %>% spread(success,Freq)
names(target_india)<-c("target","fail","success")
target_india<-target_india%>% mutate(total=fail+success,successRate=round((success/(success+fail))*100,2) , failRate=round((fail/(success+fail))*100,2))

#barplot with distribution of the most attacks in and its target
ggplot(aes(x = reorder(target,total),y=total),data=target_india) + 
  geom_bar(stat="identity",color="white",fill="#FE0202") +
  coord_flip() +
  labs(x ="Target Type ", y ="Total number of terror attacks")
#Most attacks targetted for Private citizens and property followed by police and goverment


#---------Most successful targets----------#
theme_set(theme_classic())

ggplot(aes(x = reorder(target,successRate),y=successRate),data=target_india) +
  geom_col(width=0.7,fill="#F22424",alpha=0.7) + coord_flip() +
  labs(x="Target of Terror attacks",y="% of Successul Attacks")





```



-------------------



####Time series analysis of Number of kills over the Years

I will create a dataframe consiting of year and the total number of kills for that particular year.

```{r}

#creating a new data frame-grouping by year and summarising by total sum of kills for a year
yearkillsIndia<-na.omit(indiaAttack) %>%group_by(iyear) %>% summarise(nkills=sum(nkill))

#Time series analysis
theme_set(theme_bw())
ggplot(aes(x = iyear, y =nkills),data=yearkillsIndia) +
  geom_point(color="purple",size=2) + 
  geom_line(color="#9124F2") + 
  scale_x_continuous(limits=c(1975,2016),breaks=seq(1975,2016,4)) + 
  labs(x="Year",y="number of Kills")

#highest number of kills in year 2010

#another time series chart using dygraphs
require(dygraphs)
dygraph(yearkillsIndia) %>%
  dyHighlight(highlightCircleSize = 5, 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = FALSE)





```

The above time series plot shows that the highest number of kills were in year __2010__.


------------------------------

###Group names associated with Terrorist activities

Creating a new dataframe which is grouped by the __Group name__ .

```{r,message=FALSE,warning=FALSE,fig.height=6,fig.width=12}
gnamedf<-indiaAttack %>% group_by(gname) %>%
        select(success,attacktype1_txt,targtype1_txt,nkill,iyear,provstate,city)


mostSuccessGroup<-as.data.frame(table(gnamedf$gname,gnamedf$success,
                                      gnamedf$attacktype1_txt))

#spreading the dataframe
mostSuccessGroup<-mostSuccessGroup %>% spread(Var2,Freq)
names(mostSuccessGroup)<-c("gname","attacktype","Unsuccessful","success")

SuccessGroup<-mostSuccessGroup%>% group_by(gname) %>%
          select(success,Unsuccessful,attacktype) %>%
          arrange(desc(success))
         

#considering only groups which have successful attack>20 and removing unknown groups                        
SuccessGroup <- SuccessGroup %>% filter(gname!="Unknown",success>20)

#Plotting a Barplot of Gropus vs number of successfull attacks 
theme_set(theme_bw())
p<-ggplot(aes(x=reorder(gname,success),y=success),data=SuccessGroup) + 
  geom_col(aes(fill=attacktype)) + 
  coord_flip()+
  labs(x="Group Name",y="Number of Successful attacks",title="Groups and their successful attacks",
       fill="Type of attack") + 
  scale_y_continuous(breaks=seq(0,2000,200))

p+scale_fill_brewer(palette = "Dark2")




```
 
-----------------------


### The most used Weapon and weapon types used in Attacks

Now let's analyze the most used weapons which was used by the extremists to execute the attacks.We need to create a separate data frame for arms and weapons used.


```{r}
#grouping by group_name
weapondf<-na.omit(indiaAttack) %>% filter(gname!="Unknown") %>%
  group_by(gname) %>%
  select(success,nkill,weaptype1_txt ) %>%
  arrange(desc(nkill))

#data frame with most number of kills and grouped by Group name
mostkilldf<- weapondf %>% group_by(gname) %>%
  summarise(sum_kill=sum(nkill)) %>%
  arrange(desc(sum_kill)) %>%
  top_n(20)

#plot of most kills by which terrorist group
theme_set(theme_bw())
ggplot(aes(x = reorder(gname,sum_kill) , y = sum_kill ),data = mostkilldf) +
  geom_col(fill="red") + 
  coord_flip() +
  labs(x = "Terrorist group",y="total number of kills",title="Plot of total kills vs Terrorist Group" )
  

#data frame for weapons which killed most people
weaponKill<-weapondf %>% group_by(weaptype1_txt) %>% 
  summarise(sum_kill=sum(nkill)) %>%
  filter(sum_kill >0 ) %>% 
  arrange(desc(sum_kill))
weaponKill




```

